### 2008-03-21 trunk/Papers/dl-gre-08/bisim.tex
Changed talk in Section 2 from bisimulation to simulation.

\begin{definition}
Formulas $\varphi$ of $\alc$ are generated by the following grammar:
$$
\varphi,\varphi' ::= p \mid \neg \varphi \mid \varphi \sqcap \varphi'
\mid \exists R. \varphi
$$
where $p$ is in the set of propositional symbols \prop, $R$ is in the
set of relational symbols \rel.

$\el$ is the negation-free fragment of $\alc$.

Formulas of both $\alc$ and $\el$ are interpreted in ordinary
relational first-order models $\gM = (\Delta^\gM,|\cdot|^\gM)$ where
$\Delta^\gM$ is some non-empy set, and $|\cdot|^\gM$ is an
interpretation function such that:
$$
\begin{array}{ccl}
|p|^\gM & \subseteq & \Delta^\gM  \mbox{ for $p \in \prop$}\\
|R|^\gM & \subseteq & \Delta^\gM \times \Delta^\gM  \mbox{ for $R \in \rel$}\\
|\neg \varphi|^\gM & = & \Delta^\gM \backslash |\varphi|^\gM\\
|\varphi \sqcap \varphi'|^\gM & = & |\varphi|^\gM \cap |\varphi'|^\gM\\
|\exists R.\varphi|^\gM & = & \{i \mid \mbox{for some } i', (i,i') \in |R|^\gM\\
& & \mbox{ and } i' \in |\varphi|^\gM \}.\\
\end{array}
$$
Given a model $\gM$ and an individual $i$ in the domain of $\gM$, let
$\propm(i) = \{p \in \prop \mid i \in |p|^\gM\}$.
\end{definition}

@@

\begin{definition}
Given a model $\gM = (\Delta^\gM,|\cdot|^\gM)$, two elements $i$ and
$i'$ of $\Delta^\gM$ are \emph{$\alc$-similar} if there is a relation $Z
\subseteq \Delta^\gM \times \Delta^\gM$ such that 
\begin{itemize}
\item[i)] $(i,i') \in Z$.\\[-1.5em]
\item[ii)] if $(e_1, e_2) \in Z$ then $\propm(e_1) = \propm(e_2)$.\\[-1.5em]
\item[iii)] if $(e_1,e_2) \in Z$ and for some $R \in \rel$ there is $e_1'$ such that
$(e_1,e_1') \in |R|^\gM$ then there is $e_2'$ such that $(e_2,e_2') \in |R|^\gM$ and
$(e_1',e_2') \in Z$.\\[-1.5em]
\item[iv)] if $(e_1,e_2) \in Z$ and for some $R \in \rel$ there is $e_2'$ such that
$(e_2,e_2') \in |R|^\gM$ then there is $e_1'$ such that $(e_1,e_1') \in |R|^\gM$ and
$(e_1',e_2') \in Z$.
\end{itemize}
A maximal subset of $\Delta^\gM$ all of whose elements are pairwise
bisimilar is called an \emph{\alc-bisimulation class} of $\gM$.
\end{definition}

The crucial property that makes bisimulation relevant for GRE is the
following theorem \cite{blac:moda01}.

\begin{theorem}\label{bisim}
  Given a finite model $\gM$, $i$ and $i'$ are $\alc$-bisimilar in
  $\gM$ if and only if there is no $\alc$-formula $\varphi$ such that
  $i \in |\varphi|^\gM$ and $i' \not \in |\varphi|^\gM$.
\end{theorem}

In other words, $\alc$-bisimilar individuals are indistinguishable in
the $\alc$ language.  In the example in Fig.~\ref{fig:dale-haddock}a,
the two individuals $t_1$ and $t_2$ are \alc-bisimilar; and indeed,
they satisfy exactly the same \alc-formulas.  Although we can't go
into details here, it is possible to define a suitable notion of
$\el$-bisimulation for which the analogous theorem holds
\cite{kurt:expr99}.  In general, \el-bisimulation is a weaker notion
than \alc-bisimulation; for instance, in Fig.~\ref{fig:dale-haddock},
$f_1$ and $f_2$ are \el-bisimilar \todo{oops! here's the asymmetry,
  there is a formula for $f_2$ but not for $f_1$} (no \el-formula will
distinguish them), but we have already seen that they satisfy
different \alc-formulas.

>>>>

\begin{definition}
Formulas $\varphi$ of $\alc$ are generated by the following grammar:
$$
\varphi,\varphi' ::= p \mid \neg \varphi \mid \varphi \sqcap \varphi'
\mid \exists R. \varphi
$$
where $p$ is in the set of propositional symbols \prop, $R$ is in the
set of relational symbols \rel. $\el$ is the negation-free fragment of $\alc$.

Formulas of both $\alc$ and $\el$ are interpreted in ordinary
relational first-order models $\gM = (\Delta^\gM,|\cdot|^\gM)$ where
$\Delta^\gM$ is some non-empy set, and $|\cdot|^\gM$ is an
interpretation function such that:
$$
\begin{array}{ccl}
|p|^\gM & \subseteq & \Delta^\gM  \mbox{ for $p \in \prop$}\\
|R|^\gM & \subseteq & \Delta^\gM \times \Delta^\gM  \mbox{ for $R \in \rel$}\\
|\neg \varphi|^\gM & = & \Delta^\gM \backslash |\varphi|^\gM\\
|\varphi \sqcap \varphi'|^\gM & = & |\varphi|^\gM \cap |\varphi'|^\gM\\
|\exists R.\varphi|^\gM & = & \{i \mid \mbox{for some } i', (i,i') \in |R|^\gM\\
& & \mbox{ and } i' \in |\varphi|^\gM \}.\\
\end{array}
$$

Given a model $\gM$ and an individual $i$ in $\gM$, let
$\propm(i) = \{p \in \prop \mid i \in |p|^\gM\}$.
\end{definition}

@@

\begin{definition}
Given a model $\gM = (\Delta^\gM,|\cdot|^\gM)$, and two individuals $i$ and $i'$ in $\gM$, an \emph{$\alc$-simulation} linking $i$ to $i'$  is a relation $Z
\subseteq \Delta^\gM \times \Delta^\gM$ such that 
\begin{itemize}
\item[i)] $(i,i') \in Z$.\\[-1.5em]
\item[ii)] if $(e_1, e_2) \in Z$ then $\propm(e_1) = \propm(e_2)$.\\[-1.5em]
\item[iii)] if $(e_1,e_2) \in Z$ and for some $R \in \rel$ there is $e_1'$ such that
$(e_1,e_1') \in |R|^\gM$ then there is $e_2'$ such that $(e_2,e_2') \in |R|^\gM$ and
$(e_1',e_2') \in Z$.\\[-1.5em]
\item[iv)] if $(e_1,e_2) \in Z$ and for some $R \in \rel$ there is $e_2'$ such that
$(e_2,e_2') \in |R|^\gM$ then there is $e_1'$ such that $(e_1,e_1') \in |R|^\gM$ and
$(e_1',e_2') \in Z$.
\end{itemize}
%A maximal subset of $\Delta^\gM$ all of whose elements are pairwise
%bisimilar is called an \emph{\alc-bisimulation class} of $\gM$.
\end{definition}

We will say that an individual $i$ is \emph{\alc-similar} to $i'$
in a given model $\gM$ if for any formula $\varphi \in \alc$ $i \in \interp{\varphi}$ then
$i' \in \interp{\varphi}$.  Notice that, given that the language includes negation,
the notion of $\alc$-similarity is symmetric: if $i$ is \alc-similar to $i'$ then
$i'$ is \alc-similar to $i$.  This is not the case for all description
logic.  In particular, the notion of $\el$-similarity is not symmetric.
Given a model $\gM$ and an individual $i$ in $\gM$, we will write \simm$_\alc(i)$, the similarity set, to all the elements of $\gM$ which are \alc-similar to $i$.x

The crucial property that makes simulation relevant for GRE is the
following theorem \cite{blac:moda01}.

\begin{theorem}\label{bisim}
  Given a finite model $\gM$ and two individuals $i$ and $i'$ in $\gM$, there is an $\alc$-simulation linking $i$ to $i'$ if and only if $i$ is $\alc$-similar to $i'$.
\end{theorem}

Notice that, because for $\alc$ the notion of similarity si symmetric, $\alc$-similar individuals are indistinguishable in the $\alc$ language.  In the example in Fig.~\ref{fig:dale-haddock}a,
the two individuals $t_1$ and $t_2$ are \alc-similar; and indeed,
they satisfy exactly the same \alc-formulas.

Although we can't go
into details here, it is possible to define a suitable notion of
$\el$-simulation for which the analogous theorem holds
\cite{kurt:expr99}.  In general, \el-simulation is a weaker notion
than \alc-simulation; for instance, in Fig.~\ref{fig:dale-haddock},
$f_1$ is \el-similar to $f_2$, but not vice-versa.
